Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-15 paper/main.tex
Updated paper

In section \ref{sec:sim}, we describe the simulation used to integrate the equations of motion and ins section

@@

\subsection{Fit process}
\label{sec:fit}
Given synthetic data, an Affine Invariant Markov Chain Monte Carlo (MCMC) Ensemble sampler was used to generate PPDs from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), and were bounded by $|\gamma_0| < \pi/4$, and bounds on $K_{2 m}$ given in equation \ref{eqn:parameter-bounds}. Note that $\gamma_0$ is degenerate with $\gamma_0 + \pi/2$ since this is equivalent to re-labeling $\unit y$ as $\unit x$ and $\unit x$ as $-\unit y$. The other bounds were $|K_{3m}| < 1$.

@@

@@

\subsection{Spin evolution}

In figure \ref{fig:example-data}, we present example spin data generated via our simulation. A population of one thousand asteroids with identical initial conditions except for $\gamma_0$, $K_{20}$, and $K_{22}$ were simulated on a close Earth encounter. The exact parameters used were the symmetric and asymmetric cases described in appendix \ref{app:reference-configs}. Bands containing 68.3\%, 95.5\%, and 99.7\% of the population's spin are shown, as is the spin of the reference asteroids in black.

@@

\subsection{Example fit results}

Here, we present example results of the MCMC fit described in section \ref{sec:fit}, applied to encounter data of the asymmetric reference asteroid. The fit results were consistent with the true density moments, and produced consistent data.

Figure \ref{fig:example-residuals} shows the spin data with the best-fit overlaid in the top pane and the residuals in the bottom pane. Uncertainties are plotted on the residuals corresponding to the square root of the diagonal entries of the covariance matrix (correlations not included).

Figure \ref{fig:example-corner} shows a corner plot of the ten fitted parameters' posterior distributions, marginalized to functions of one or two variables. The true parameters are shown as lines. Note that the true parameters usually lie within 1- or 2$\sigma$ of the $\Delta K_{\ell m} = 0$, where $\Delta K_{\ell m}$ is the difference between the posterior $K_{\ell m}$ and the true $K_{\ell m}$.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figs/example-corner.pdf}
  \caption{A corner plot for the fit results ten parameters to synthetic data simulated for an asymmetric reference asteroid. Marginal posterior PDFs for each parameter are shown (histograms) along with two-dimensional PDFs (contours). Individual points in the contours are samples from the MCMC fit, and the contours enclose 1-, 2-, and 3-$\sigma$ of them. True values (which are zero here due to the parameters being displayed relative to the true values) are also shown (blue lines).}
  \label{fig:example-corner}
\end{figure*}

\subsection{Density stuff?}

% \afterpage{
%   \begin{landscape}
%     \begin{figure}
%     \centering
%     \includegraphics[height=0.89\textheight]{figs/scan-period.pdf}
%     \caption{caption}
%     \label{fig:label}
%     \end{figure}
%   \end{landscape}
% }
% PLOTS THAT COULD BE HERE
% rp
% vp

In this section, we assess the sensitivity of the PPDs for the first-order parameters ($\gamma_0$, $K_{20}$, and $K_{22}$), and the second-order parameters ($K_{3m}$) to many possible close encounter configurations. Specifically, we test sensitivity to the encounter's orbital parameters, the degree of observational uncertainty, the asteroid's radius, the cadence of observation, the amount of data collected at perigee, the direction of the initial spin pole, the asteroid period, and the central body oblateness. These effects are presented roughly in order of how much they affect the parameter precision, from highest to lowest.

Parameter precision of 1$\sigma$ is defined such that 68.27\% of the PPD lies within 1$\sigma$ of the mean of the PPD. 2$\sigma$ is defined likewise for 95.45\% of the PPD. The PPDs are usually roughly Gaussian, so that the upper and lower error bars are usually roughly equal and 1$\sigma$ is approximately the standard deviation of the PPD.

In all cases, we use the configuration of the asymmetric reference asteroid (appendix \ref{app:reference-configs}) unless otherwise stated. We further present a test of how precision depends on cadence, asteroid period, and the duration of the encounter in appendix \ref{app:cadence-tests}, and we compare precision for a Jupiter and an Earth encounter in appendix \ref{app:jupiter-earth}.

\subsection{Orbital elements}
\label{sec:scan-orbit}
A Keplerian orbit is completely described by five parameters, but three describe the orbit's orientation with respect to the central body. The orbit can be rotated to the $\unit X \unit Y$-plane, as mandated by our coordinate definitions, by changing the density moments of the central body. Since $J_{00}$ is unchanged by this rotation and $J_{1m}=0$, the orbit orientation is irrelevant until the $J_{2m}$ terms of equation \ref{eqn:tidal-torque}, and we do not investigate them here.

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-vex.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of orbital excess velocity $v_\infty$. The vertical dashed line indicates the reference asteroid value of 6 km s$^{-1}$. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-vex}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-perigee.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of perigee distance $r_p$. The vertical dashed line indicates the reference asteroid value of 5 Earth radii. The dotted curve indicates a power-law fit (see text). The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-perigee}
\end{figure}

@@

\subsection{Observational Uncertainty}
\label{sec:scan-uncertainty}
There are two parameters, $\sigma_\theta$ and $\sigma_\rho$, which govern the observational uncertainty of the data set (defined in section \ref{sec:uncertainty}). Rather than explore the full space spanned by these two values, we measure how parameter uncertainties depend on the product of uncertainties $\sigma_\theta\sigma_\rho$ (with the radio fixed), and the ratio $\sigma_\rho / \sigma_\theta$ (with the product fixed).

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-product.pdf}
  \caption{1 and 2$\sigma$ confidence intervals divided by $\sigma_\theta$ for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty product $\sigma_\theta \sigma_\rho$. The vertical dashed line indicates the reference asteroid value of $10^{-9}$. The red vertical lines indicate when $\sigma =0.01$.}
  \label{fig:scan-product}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-ratio.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty ratios $\sigma_\rho / \sigma_\theta$. The reference asteroid value is $\sigma_\rho/ \sigma_\theta =10^{-5}$. The red vertical lines indicate when $\sigma = 0.01$. (Lines are not shown for $K_{3m}$, $m < 3$ because they coincide with the vertical axis.)}
  \label{fig:scan-ratio}
\end{figure}

@@

\subsection{Asteroid shape}
\label{sec:scan-shape}

@@

\begin{figure*}
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-1-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-2-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-3-sigma.pdf}

  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-1-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-a-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-b-sigma.pdf}

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-am.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of asteroid length $a_\mathcal{A}$. The reference cadence is 2 minutes. The red vertical lines indicate when $\sigma = 0.01$. The black dashed line indicates the reference value of $a_\mathcal{A}=1000$ m.}
  \label{fig:scan-am}
\end{figure}

@@

\subsection{Cadence}
\label{sec:scan-cadence}

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-cadence.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational cadence $\Delta t$ The reference cadence is 2 minutes. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-cadence}
\end{figure}

@@

\subsection{Perigee gap}
\label{sec:scan-gap}
In certain circumstances, spin data might not be able to be captured for a close encounter at perigee. The asteroid might dip below the horizon, or it might pass too close to the sun to be observed. Generally, angular velocity data can be collected when the asteroid is distant from the central body, where torque is low. There, the angular velocity evolution is dominated by torque-free precession dictated by the moment of inertia components. That zero-torque data can still be used to fix $K_{20}$ and $K_{22}$ as in \cite{MOSKOVITZ2020113519}. However, $K_{3m}$ are not extractable from precession data alone. We are therefore curious as to how our posterior uncertainties change due to lack of data during the encounter perigee.

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/observation-gap.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of a data gap of width $T$ at perigee. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:observation-gap}
\end{figure}

@@

\subsection{Initial spin pole}
\label{sec:scan-spin}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figs/spin-pole.pdf}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.45\textwidth]{figs/spin-pole-avg-mollweide.pdf}
  \includegraphics[width=0.45\textwidth]{figs/spin-pole-avg-polar.pdf}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}

@@

\subsection{Rotational period}
\label{sec:scan-period}

@@

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-period.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of initial rotational period $P_\omega$. The reference value of 9 hr is shown as a vertical dotted line. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-period}
\end{figure}

@@

\subsection{Central body oblateness}
\label{sec:scan-oblateness}

@@

\section{Density distributions}
\label{sec:distros}

Hitherto, we have only discussed the density moments $K_{\ell m}$ of the asteroid rather than the true density distribution $\rho(\bm r)$. This is because only the density moments are observable by tidal torque interactions. However, by making sufficient assumptions about the density distribution, we can nevertheless estimate $\rho(\bm r)$ from $K_{\ell m}$.

In this section, we outline three different assumptions which yield different types of density distributions for the same $K_{\ell m}$. One of them can be chosen, or a fourth created, depending on what features are desired to be represented in the density distribution. All three of these models assume that $a_\mathcal{A}$ and the asteroid shape is known (for example, by light curve or radar data). Uncertainty on this shape estimate is assumed to be small. The models then produce a density distribution $\rho(\bm r)$ with uncertainty from $K_{\ell m}$ propagated to uncertainty in $\rho$.

In section \ref{sec:results}, we were forced to fix $a_\mathcal{A}$ rather than fit for it because $a_\mathcal{A}$ is degenerate with scaling the $K_{3m}$ components. However, $a_\mathcal{A}$ contains information about the density distribution as shown by its definition in equation \ref{eqn:am}. It might therefore be unreasonable to assume that it is exactly determined by light-curve or radar analysis. If $a_\mathcal{A}$ is too uncertain to be used as a known quantity when extracting density distributions from the moments, then the following models are flexible enough that the $a_\mathcal{A}$ constraint can be removed.

In section \ref{sec:general-density}, we describe the general form our three models must take. Then in sections \ref{sec:likelihood} to \ref{sec:lumpy} we lay out each individually and compare them in section \ref{sec:density-compare}. In appendix \ref{app:find-surface}, we introduce a new model which does not assume that the asteroid shape is known, instead assuming that the density is constant. This allows us to compute the asteroid shape from $K_{\ell m}$.

\subsection{General density model design}
\label{sec:general-density}

\subsection{The ``likelihood'' model}
\label{sec:likelihood}

\subsection{The ``harmonic'' model}
\label{sec:harmonic}

\subsection{The ``lumpy'' model}
\label{sec:lumpy}

\subsection{Comparisons between density models.}
\label{sec:density-compare}

To test the properties of the three density models defined above, we simulate several asteroids with different shapes and density distributions on a close Earth encounter with the reference asteroid orbital and observational parameters. Density moments are extracted via our fit process and density distributions extracted by the three above methods. Below, we compare the resulting distributions to understand the performance of the extraction methods.

@@

  \includegraphics[width=0.33\textwidth]{figs/tet-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/tet-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/tet-lumpy.pdf}
  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density asteroid. The asteroid shapes are, from top to bottom, the reference asymmetric ellipsoid, the reference symmetric asteroid, the dumbbell model, and a tetrahedron. Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:uniform-density}
\end{figure*}

@@

\include{appendices}

>>>>

In section \ref{sec:sim}, we describe the simulation used to integrate the equations of motion. In section \ref{sec:uncertainty}, we define the uncertainty model used to add noise to the data, and we describe how PPDs for density moments were extracted from spin-pole data in section \ref{sec:fit}. Finally, we outline a general model to extract density distributions from density moments in section \ref{sec:general-density} and define specific model implementations in section \ref{sec:specific-density}

@@

\subsection{Extracting density moments from spin data}
\label{sec:fit}
Given synthetic data, an Affine Invariant Markov Chain Monte Carlo (MCMC) Ensemble sampler was used to generate PPDs from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), and were bounded by $|\gamma_0| < \pi/4$, and bounds on $K_{2 m}$ given in equation \ref{eqn:parameter-bounds}. Note that $\gamma_0$ is degenerate with $\gamma_0 + \pi/2$ since this is equivalent to re-labeling $\unit y$ as $\unit x$ and $\unit x$ as $-\unit y$. The other bounds were $|K_{3m}| < 1$.

@@

\subsection{Extracting density distributions from density moments: general model design}
\label{sec:general-density}

The asteroid density distribution $\rho_\mathcal{A}(\bm r)$ is not uniquely constrained via tidal torque interactions because only the density moments $K_{\ell m}$ contribute to equation \ref{eqn:tidal-torque}. However, by making sufficient assumptions about the density distribution, we can nevertheless estimate $\rho(\bm r)$ from $K_{\ell m}$. Here, we outline three different assumptions which constitute three models for extracting density distributions from density moments. One of them can be chosen, or another created, depending on what features are desired to be represented in the density distribution.

All three of these models assume that $a_\mathcal{A}$ and the asteroid shape is known (for example, by light curve or radar data). Uncertainty on this shape estimate is assumed to be small. The models then produce a density distribution $\rho_\mathcal{A}(\bm r)$ with uncertainty from $K_{\ell m}$ propagated to uncertainty in $\rho_\mathcal{A}$. In section \ref{sec:results}, we were forced to fix $a_\mathcal{A}$ rather than fit for it because $a_\mathcal{A}$ is degenerate with scaling the $K_{3m}$ components. However, $a_\mathcal{A}$ contains information about the density distribution as shown by its definition in equation \ref{eqn:am}. It might therefore be unreasonable to assume that it is exactly determined by light-curve or radar analysis. If $a_\mathcal{A}$ is too uncertain to be used as a known quantity when extracting density distributions from the moments, then the following models are flexible enough that the $a_\mathcal{A}$ constraint can be removed. Likewise, if the asteroid shape is not known with certainty, a fourth model described in appendix \ref{app:find-surface} can be used to extract asteroid shape assuming constant density.

\subsection{Model implementations for extracting density distributions from density moments}
\label{sec:specific-density}
In the following sections, three possible models for extracting density distributions from moments (i.e., choices of $\bm B(\bm r)$ and $A$ defined above) are given.

\subsubsection{The ``likelihood'' model}
\label{sec:likelihood}

\subsubsection{The ``harmonic'' model}
\label{sec:harmonic}

\subsubsection{The ``lumpy'' model}
\label{sec:lumpy}

@@

\subsection{Density moment extraction}
To demonstrate the success of our fit process, we present example results of the MCMC fit described in section \ref{sec:fit}, applied to encounter data of the asymmetric reference asteroid. The fit results were consistent with the true density moments, and produced consistent data.

Figure \ref{fig:example-residuals} shows the spin data with uncertainty added. Data simulated given the best-fitting density moments (the mean of the PPDs) is overlaid in the top pane and the residuals are shown the bottom pane. Uncertainties are plotted on the residuals corresponding to the square root of the diagonal entries of the covariance matrix (correlations not included). The fit results are clearly consistent with the data.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figs/example-corner.pdf}
  \caption{A corner plot for the fit results ten parameters to synthetic data simulated for an asymmetric reference asteroid. Marginal posterior PDFs for each parameter are shown (histograms) along with two-dimensional PDFs (contours). Individual points in the contours are samples from the MCMC fit, and the contours enclose 1, 2, and 3$\sigma$ of them. True values (which are zero here due to the parameters being displayed relative to the true values) are also shown (blue lines).}
  \label{fig:example-corner}
\end{figure*}

Figure \ref{fig:example-corner} shows a corner plot of the ten fitted PPD, marginalized to functions of one or two variables. The true parameters are shown as lines. Note that the true parameters usually lie within 1 or 2$\sigma$ of the $\Delta K_{\ell m} = 0$, where $\Delta K_{\ell m}$ is the difference between the posterior $K_{\ell m}$ and the true $K_{\ell m}$. The PPDs are generally Gaussian and no degeneracy between fit components is visible. We performed 48 independent minimizations of the likelihood before the MCMC fit began, each with an initial point chosen randomly in the parameter space. All converged to the same minimum, further showing that the model lacks degeneracy.

\subsection{Density distribution extraction}
\label{sec:asym-density}
To demonstrate the success of our three density distribution extraction models, we show density distributions and uncertainties extracted for the asymmetric reference asteroid in figure \ref{fig:uniform-density-results}. These distributions were made given the density moment PPDs extracted in the previous section, with $a_\mathcal{A}$ fixed. The surface is fixed as a triaxial ellipsoid with semi-axis lengths such that the true values of $K_{\ell m}$ are consistent with uniform density (equation \ref{eqn:ellipsoid-axes}).

  \includegraphics[width=0.33\textwidth]{figs/asym-ell-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-unc-harmonic.pdf}\hfill\hspace{0.33\textwidth}

  \caption{Density distributions (\textit{top}) and uncertainties (\textit{bottom}) extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density, asymmetric reference asteroid. Density is normalized so that mean density is one.}
  \label{fig:uniform-density-results}
\end{figure*}

Also in each pane of figure \ref{fig:uniform-density-results} is $\Delta K_{\ell m}$, the squared magnitude of difference between the fitted density moments and the density moments of the final distribution. This difference is non-zero due to numerical error in the likelihood model case, but for the harmonic and the lumpy models, they express the degree to which the extracted moments cannot be represented with harmonic or lumpy distributions respectively.

The likelihood and harmonic models produce mostly the same density distributions. This is generally true for uniform density asteroids, which have harmonic density distributions. Consequentially, $\Delta K_{\ell m}$ is low for the harmonic model. This similarity lends confidence that the models produced are truly representative of the asteroid since they are produced via two independent methods.

The distributions' non-uniformity is limited for the harmonic and likelihood models. This non-uniformity is driven by error in the estimates of the density moments. Note that most of the asteroid has density close to one in both cases, with densities of 40\% or 160\% of the mean attained only in small regions. The lumpy model yields a much more uniform distribution because it is designed to produce uniform lumps. However, it does not produce a density distribution consistent with those extracted from spin data (i.e., $\Delta K_{\ell m}$ is usually large). This is because the lump found extends outside the asteroid, violating the assumptions of the model --- a sign that the lump found does not really exist inside the asteroid.

The size of uncertainty is such that the density distribution is consistent with uniform at most points for the harmonic and likelihood models (within 1-2$\sigma$), with a small number of low-uncertainty regions that are less consistent with uniform. Uncertainty is generally larger towards the edges of the distribution, both for these shapes and for the shapes shown in the following sections.

In this section, we discuss the physical properties of the asteroid encounter system (section \ref{sec:spin-evolution}), as well as the sensitivity of the density moment uncertainty to encounter properties (section \ref{sec:fit-uncertainty}), together with an analysis of advantages and disadvantages Jupiter encounters over Earth (section \ref{sec:jupiter-earth})Finally, we discuss the kinds of distributions produced by the density distribution extraction methods (section \ref{sec:density-compare}).

\subsection{Spin evolution}
\label{sec:spin-evolution}
In figure \ref{fig:example-data}, we present example spin data generated via our simulation. A population of one thousand asteroids with identical initial conditions except for $\gamma_0$, $K_{20}$, and $K_{22}$ were simulated on a close Earth encounter. The exact parameters used were the symmetric and asymmetric cases described in appendix \ref{app:reference-configs}. Bands containing 68.3\%, 95.5\%, and 99.7\% of the population's spin are shown, as is the spin of the reference asteroids in black.

@@

\subsection{Dependence of parameter uncertainty on encounter parameters}
\label{sec:fit-uncertainty}
Parameter precision of 1$\sigma$ is defined such that 68.27\% of the PPD lies within 1$\sigma$ of the mean of the PPD. 2$\sigma$ is defined likewise for 95.45\% of the PPD. The PPDs are usually roughly Gaussian, so that the upper and lower error bars are usually roughly equal and 1$\sigma$ is approximately the standard deviation of the PPD. In all cases, we use the configuration of the asymmetric reference asteroid (appendix \ref{app:reference-configs}) unless otherwise stated.

\subsubsection{Orbital elements}
\label{sec:scan-orbit}
A Keplerian orbit is completely described by five parameters, but three describe the orbit's orientation with respect to the central body. The orbit can be rotated to the $\unit X \unit Y$-plane, as mandated by our coordinate definitions, by changing the density moments of the central body. Since $J_{00}$ is unchanged by this rotation and $J_{1m}=0$, the orbit orientation is irrelevant until the $J_{2m}$ terms of equation \ref{eqn:tidal-torque}, and we do not investigate them here.

@@

@@

@@

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all-1.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-perigee}
    \label{fig:scan-vex}
    \label{fig:scan-theta}
    \label{fig:scan-rho}
    \end{figure}
  \end{landscape}
}

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all-2.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of various encounter parameters (\textit{columns}). The vertical dashed line indicates the reference asteroid values of the parameters. The red vertical lines indicate when $\sigma =0.01$.}
    \label{fig:scan-am}
    \label{fig:scan-cadence}
    \label{fig:scan-period}
    \label{fig:scan-gap}
    \end{figure}
  \end{landscape}
}

\subsubsection{Observational Uncertainty}
\label{sec:scan-uncertainty}
There are two parameters, $\sigma_\theta$ and $\sigma_\rho$, which govern the observational uncertainty of the data set (defined in section \ref{sec:uncertainty}). Rather than explore the full space spanned by these two values, we measure how parameter uncertainties depend on the product of uncertainties $\sigma_\theta\sigma_\rho$ (with the radio fixed), and the ratio $\sigma_\rho / \sigma_\theta$ (with the product fixed).

@@

@@

@@

\subsubsection{Asteroid shape}
\label{sec:scan-shape}

@@

\begin{figure*}
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-1-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-2-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-theta-3-sigma.pdf}

  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-1-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-a-sigma.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/probe-space-ab-b-sigma.pdf}

@@

@@

\subsubsection{Cadence}
\label{sec:scan-cadence}

@@

@@

\subsubsection{Perigee gap}
\label{sec:scan-gap}
In certain circumstances, spin data might not be able to be captured for a close encounter at perigee. The asteroid might dip below the horizon, or it might pass too close to the sun to be observed. Generally, angular velocity data can be collected when the asteroid is distant from the central body, where torque is low. There, the angular velocity evolution is dominated by torque-free precession dictated by the moment of inertia components. That zero-torque data can still be used to fix $K_{20}$ and $K_{22}$ as in \cite{MOSKOVITZ2020113519}. However, $K_{3m}$ are not extractable from precession data alone. We are therefore curious as to how our posterior uncertainties change due to lack of data during the encounter perigee.

@@

@@

\subsubsection{Initial spin pole}
\label{sec:scan-spin}

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/spin-pole.pdf}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-mollweide.pdf}
  \includegraphics[width=0.3\textwidth]{figs/spin-pole-avg-polar.pdf}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}

@@

\subsubsection{Rotational period}
\label{sec:scan-period}

@@

@@

\subsubsection{Central body oblateness}
\label{sec:scan-oblateness}

@@

\subsection{Comparison of Jupiter and Earth encounters}
\label{app:jupiter-earth}

If sufficiently accurate spin pole data can be detected for non-Earth encounters, it may be possible to extract density moments for encounters with larger planets. In this appendix, we run our reference asteroid through a Jupiter encounter to analyze the differences in uncertainty.

The physical parameters of the asteroid body are kept the same as the Earth encounter case (listed in appendix \ref{app:reference-configs}), as are the observational uncertainty and cadence. The orbit is adjusted for the Jupiter case by setting a perijove distance of $r_p=5$ Jupiter radii (compared to perigee radius $r_p=5$ Earth radii for the Earth encounter). We perform two fits, each with a different value of $v_\infty$: in one case we keep the Earth-encounter value of $v_\infty = 6$ km s$^{-1}$, and in the other case we increase the velocity such that the eccentricity (and therefore the angle between the orbits' hyperbolic asymptotes) are the same for both encounters. The latter requires $v_\infty = \sqrt{G\mu_\mathcal{B}/r_p (e-1)}=32$ km s$^{-1}$. The rationale for these excess velocity values is that, for $v_\infty = 6$ km s$^{-1}$, we keep a physically typical encounter excess velocity at the expense of changing the orbit shape, whereas for $v_\infty =32$ km s$^{-1}$ the orbit shape is unchanged, although the asteroid speed is dramatically increased.

Section \ref{sec:scan-orbit} found that $v_\infty$ does not affect the posterior uncertainties, and indeed we find very little difference between the two $v_\infty$ cases. We therefore do not distinguish between the two cases further. The ratio between the posterior uncertainties in the Jupiter and the Earth encounters are shown in table \ref{tab:jupiter-uncertainty}. In all cases, the Jupiter posteriors are more uncertain than Earth posteriors.

\begin{table}
  \centering
  \begin{tabular}{c|cc}
    \hline 
    $K_{\ell m}$ & $\sigma_\text{Jupiter}/\sigma_\text{Earth}$\\ \hline 
    $\gamma_0$ & 1.6 \\
    $K_{22}$ & 2.3 \\
    $K_{20}$ & 11 \\
    $\Re K_{33}$ & 18 \\
    $\Im K_{33}$ & 18 \\
    $\Re K_{32}$ & 18 \\
    $\Im K_{32}$ & 18 \\
    $\Re K_{31}$ & 25 \\
    $\Im K_{31}$ & 10 \\
    $K_{30}$ & 53 \\ \hline
  \end{tabular}
  \caption{Ratio of posterior uncertainty for all density moments $K_{\ell m}$ between an Earth encounter and a Jupiter encounter with identical properties except for an increased perigee.}
  \label{tab:jupiter-uncertainty}
\end{table}

These uncertainty ratios can be understood as follows. The leading order of tidal torque is proportional to $\mu_\mathcal{A} / D^3$. If $D/a_\mathcal{B}$ (the ratio of the encounter distance to the central body radius) is roughly constant (as in this case, where $r_p/a_\mathcal{B}=5$), then $\mu_\mathcal{A} / D^3 \propto \rho_\mathcal{B}$ where $\rho_\mathcal{B}$ is the density of the central body. Therefore, little advantage is to be gained by looking for encounters of a massive planet in this sense. Since Jupiter is less dense than Earth, we expect that uncertainty in the first-order parameters $\gamma_0$ and $K_{2m}$ would be slightly worse than in the case of Earth, which is seen in table \ref{tab:jupiter-uncertainty}.

The second-order terms are damped by an additional factor of $a_\mathcal{A}/D$, which decreases if a massive central body is used. Since Jupiter is about 10 times larger in radius than Earth, we expect that the $K_{3m}$ terms are about ten times more uncertain than the $K_{2m}$ components, which is the case.

The $K_{\ell 0}$ components differ in that $\sigma_\text{Jupiter}/\sigma_\text{Earth}$ is about five times greater for $K_{\ell 0}$ than other moments of the same $\ell$. In fact, $K_{30}$ essentially fills the prior. The special properties of $K_{\ell 0}$ have been analysed noted before, in that they are always more uncertain than other moments and are particularly uncertain in regimes where the asteroid does not tumble after the perigee. In this case, the Jupiter encounter resulted in less tumbling than the Earth encounter, so the larger increase in uncertainty in $K_{\ell 0}$ shown in table \ref{tab:jupiter-uncertainty} is expected.

There are additional effects of central body mass which are not captured in this analysis. For example, encounters with massive planets are more plentiful, so that observation for a fixed period of time will lead to a larger number of observed encounters conducive to low-uncertainty moment extraction (large $a_\mathcal{A}$, small $r_p$, etc.). The distribution of $r_p$ in this encounter sample will also change; the ratio of the orbit impact parameter (the distance between the orbit asymptotes and the central body) to perigee distance is 
\begin{equation}
  \frac{b}{r_p} = \sqrt{1+2\frac{G\mu_\mathcal{B}}{r_p v_\infty^2}}.
\end{equation}
It was mentioned above that the observable perigee distance $r_p$ increases with $a_\mathcal{B} \sim \mu_\mathcal{B}^{1/3}$, so that $\frac{b}{r_p}$ is larger for more massive planets and therefore the same impact parameter leads to lower periapse. Other effects, such as a change in the physical properties of the encountering asteroids, may also affect the fit uncertainties.

Table \ref{tab:jupiter-uncertainty} shows that an encounter with Jupiter yields less precise posteriors than an encounter with Earth, all else being equal. However, these selection biases will cause more large-$a_\mathcal{A}$ and small-$r_p$ asteroids to encounter Jupiter, and therefore yield increased precision. Which of these contradicting effects dominates is not immediately clear, and depends on the specific asteroid population near Jupiter.

\subsection{Comparisons between density models}
\label{sec:density-compare}

To test the properties of the three density models defined in section \ref{sec:specific-density}, we simulate several asteroids with different shapes and density distributions on a close Earth encounter with the reference asteroid orbital and observational parameters. Density moments are extracted via our fit process and density distributions extracted by the three above methods. Below, we compare the resulting distributions to understand the performance of the extraction methods.

@@

  \includegraphics[width=0.33\textwidth]{figs/tet-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/tet-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/tet-lumpy.pdf}
  \caption{Density distributions extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density asteroid. The asteroid shapes are, from top to bottom, the reference symmetric asteroid, the dumbbell model, and a tetrahedron. Density is normalized so that mean density is one. The same distance scale is used in all figures.}
  \label{fig:uniform-density}
\end{figure*}

@@

%\include{appendices}
